---
title: "Importing Data into R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CSV

CSVs are one of the most common ways to bring data into R. R makes it easy by including a base function **read.csv()**.

The basic structure of the function is simply *read.csv(filepath)*. Be sure to save this function to a variable in R using **<-**.

```{r read csv, eval=FALSE}
# Read in the bpda addresses csv
bpda <- read.csv('bpda_addresses.csv')

# View to first 6 rows of bpda
head(bpda)
```

Add a zero at the beginning of the ZIP column using string padding. This is a neat trick you may need to use since we live in Boston and frequently zeros get dropped at the beginning of our zip codes if we're not careful.

```{r fix zip, eval=FALSE}
# Install and load stringi package
install.packages('stringi')
library(stringi)

# Pad the ZIP column with a 0 up to the fifth digit
bpda$ZIP <- stri_pad(bpda$ZIP, 5, pad = '0')

# View the same rows as before, notice ZIP now correctly includes the 0 at the beginning and is no longer an integer
head(bpda)
```

You can then write data back to csv using **write.csv**. The basic function structure is *write.csv(table, filepath)*.

```{r write csv, eval=FALSE}
# Write bpda back to csv
write.csv(bpda, 'bpda2.csv', row.names = FALSE)

```


## Google Sheets

You can access google sheets and import them as tables in R using the **googlesheets** package. This also allows you to upload to Google Sheets.

This method involves manually authenticating your google account in a browser so it is less useful than setting up a password/username but is more secure for sharing online.

```{r googlesheets, eval=FALSE}
# Install and load googlesheeets package and tidyverse package
install.packages('googlesheets') 
install.packages('tidyverse')

library(googlesheets)
library(tidyverse)

# Set up Google Sheet authorization
gs_auth(new_user = TRUE)

# Check available sheets
gs_ls()
available <- gs_ls()

# View available sheets in console
View(available)

# Read in Stairs Tracker to see how good James is at climbing stairs
stairs.title <- gs_title('Stairs Tracker')
stairs <- gs_read(stairs.title)
view(stairs)
```

Now we want to only keep records that were made by James since he is the wunderkind of Competitive Stair Climbing and everyone else pales in comparison. We can then export this table to CSV and upload it to Google Sheets.

```{r remove plebs and upload, eval=FALSE}
# Keep only rows that contain James and put them in new table
stairs.wunderkind <- stairs [ stairs$Name == 'James', ]

# Write to a CSV as you did before
write.csv(stairs.wunderkind, 'stairs_wunderkind.csv', row.names = FALSE)

# Upload this CSV to a Google Sheet
gs_upload('stairs_wunderkind.csv',
          sheet_title = 'Stairs Tracker for the Only Person Who Uses It')

```

## SQL Server Access

R can also access SQL servers to import tables using several packages. Today we'll be using **RODBC** to do this. 

You'll need to have Windows Authentication access to the database on the computer you're using to make this specific method work.

```{r sql, eval=FALSE}
# Install and load RODBC
install.packages('RODBC')
library(RODBC)

# Set up the connection
con <- odbcDriverConnect('driver={SQL Server}; server=vsql25; database=EDW_Generic; trusted_connection=true')

# Write the query
query <- "SELECT * FROM dbo.DOIT_CraigslistData WHERE bedrooms = '1BR'"

# Utilize the query
craigslist.1br <- sqlQuery(con, query)
view(craigslist.1br)

```





















